---
title: "Forecast evaluation - Analysis and plots"
output: html_notebook
---

This notebook analyses the performance of unconditional and conditional forecasts for a large number of macroeconomic time series. 

## Set-up

```{r}
setwd("C:/Users/Philipp/Documents/GitHub/condfcast-precsampler/eval")
library(ggplot2)
library(tidyr)
library(dplyr)
library(lubridate)
library(ggsci)
library(latex2exp)
```

## Load model and benchmark forecasts

```{r}
# models
load("df_eval_models.Rda")
df_eval_models$quarter <- as_date(df_eval_models$quarter)
df_eval_models$vintage <- as_date(df_eval_models$vintage)

# benchmarks
load("benchmarks/df_eval_benchmark.Rda")
```

## Manually remove some series and vintages

For **gross value added in finance, rent and professional servies** there are only vintages up until 2012, after which disaggregate series for the different sector are available. I therefore do not consider the series

```{r}
df_eval_models <- filter(df_eval_models, !(mnemonic == "gva_freprof"))
```


For two labor market series, there are some missing vintages. These lead to an information set that is inconsistent with the rest of the evaluation period. For example, because of the missing vintages and data, a calendar-based forecast horizon of $-1$, say, does not reflect the same horizon in terms of observations. 

The series are

- `h_ind`: For hours worked in the manufacturing sector, several vintages in the year 2009 are missing. Specifically, no vintages are published between March 16 2009 and March 4 2010. However, it seems that the usual release pattern is only followed from March 15 2010 onwards. On March 4, only the values up to December 2009 are published. Therefore, all vintages between March 16 2009 and March 15 2010 are excluded from the analysis

```{r}
df_eval_models <- filter(df_eval_models, 
                        !(mnemonic == "h_ind" &
                              vintage < "2010-03-15" &
                                vintage > "2009-03-16")
            )

df_eval_benchmark <- filter(df_eval_benchmark, 
                           !(series == "h_ind" & 
                               vintage < "2010-03-15" & 
                                  vintage > "2009-03-16")
                          )
```


- `emp`: For total employment, the "December 2013" vintage was released in early January 2014 (quite likely due to the original or usual publication date falling on a bank holiday). The previous vintage was published on November 28. Any vintages in between those two dates are removed! 

```{r}
df_eval_models <- filter(df_eval_models, 
                        !(mnemonic == "emp" &
                              vintage < "2014-01-07" &
                                vintage > "2013-11-28")
                  )

df_eval_benchmark <- filter(df_eval_benchmark, 
                           !(series == "emp" &
                                vintage < "2014-01-07" &
                                  vintage > "2013-11-28")
                          )
```

## Function to calculate relative RMSFE, average log score, CRPS

Function that calculates the forecast accuracy of a selected **model** in a selected **subsample** relative to the benchmark. Output is dataframe `df_eval`
```{r}
calc_relative_forecast_accuracy <- function(df_eval_models, df_eval_benchmark, str_model, subsample)
{
  # calculate RMSFE, log score and CRPS over subsample for chosen model and benchmark
  df_eval_models %>% 
  filter(model == str_model,
         quarter >= subsample[1],
         quarter <= subsample[2]) %>%
  group_by(mnemonic, horizon, type, model) %>%
  summarise(rmsfe = sqrt(mean(sfe)), 
            mean_logs = mean(logs),
            mean_crps = mean(crps)) %>%
  ungroup() -> df_eval_models_avg

  df_eval_benchmark %>% 
  filter(quarter >= subsample[1],
         quarter <= subsample[2]) %>%
  group_by(series, horizon) %>%
  summarise(rmsfe = sqrt(mean(sfe)), 
            mean_logs = mean(logs),
            mean_crps = mean(crps)) %>%
  ungroup() -> df_eval_benchmark_avg
  
  # relative RMSFE, log score and CRPS
  df_eval_models_avg %>% 
    merge(select(df_eval_benchmark_avg, 
                  horizon,
                  mnemonic = series, 
                  rmsfe_ar = rmsfe,
                  logs_ar = mean_logs, 
                  crps_ar = mean_crps
                ), 
          by = c("horizon", "mnemonic")) %>%
    mutate(rel_rmsfe = rmsfe / rmsfe_ar,
           rel_logs = mean_logs / logs_ar,
           rel_crps = mean_crps / crps_ar) -> df_eval
  
  # merge with categories
  source("../data/realtime_data.R")
  tmp <- realtime_data()
  tmp <- select(tmp, mnemonic, category)
  df_eval <- merge(df_eval, tmp, by = "mnemonic")
  rm(tmp)
  
  # select columns
  df_eval <- select(df_eval, mnemonic, horizon, type, category, rel_rmsfe, rel_logs, rel_crps)
  
  return(df_eval)
}
```


## Analysis

```{r}
horizon_as_factor <- function(x)
{
  return(factor(paste0("h=", x)))
}
```



```{r}
str_model <- "Nr2_Nj1_Np2_Ns0"
subsample <- c(min(df_eval_models$quarter), max(df_eval_models$quarter))

df_eval <- calc_relative_forecast_accuracy(df_eval_models, 
                                           df_eval_benchmark, 
                                           str_model, 
                                           subsample)

```
### Point forecast accuracy

```{r}
# df_plot
str_title <- "rel. RMSFE"
df_eval %>%
  filter(horizon >= 0) %>%
  pivot_longer(c(rel_rmsfe, rel_crps, rel_logs), names_to = "score", values_to = "value") %>%
  filter(value != Inf, score == "rel_rmsfe") %>%
  pivot_wider(names_from = type, values_from = value) -> df_plot 

# data-dependent lims!
ylims <- c(min(min(df_plot$unconditional), min(df_plot$conditional_hard)),
           max(max(df_plot$unconditional), max(df_plot$conditional_hard)))
xlims <- ylims

# manually set lims, excluding "outliers"
# ylims <- c(0.8, 1.2)
# xlims <- ylims

ggplot(df_plot, aes(x = unconditional, 
                    y = conditional_hard, 
                    color = category)
       )+
  geom_point(size = 2, alpha = 0.7)+
  geom_abline()+
  geom_vline(xintercept = 0,  size = 0.2)+
  geom_hline(yintercept = 0, size = 0.2)+
  facet_wrap(~horizon_as_factor(horizon), nrow = 1)+
  xlim(xlims)+
  ylim(ylims)+
  scale_color_jco()+
  #theme_minimal()+
  theme(legend.position="bottom", 
        legend.text = element_text(size = 8),
        legend.title = element_blank())+
  labs(title = "", x = "unconditional forecasts", y = "conditional forecasts")
```

#### Export plot to pdf

```{r}
ggsave(filename = "../figures/fig_eval_rmsfe.pdf", 
       width = 10, height = 6, units = "in")
```


#### Analysis


```{r}
df_plot %>% 
  filter(horizon == 0) %>% 
  select(mnemonic, unconditional, horizon) %>% 
  pivot_wider(names_from = "horizon", values_from = "unconditional") %>% 
  arrange(`0`) %>% 
  head(10)

df_plot %>% 
  filter(horizon == 1) %>% 
  select(mnemonic, unconditional, horizon) %>% 
  pivot_wider(names_from = "horizon", values_from = "unconditional") %>% 
  arrange(`1`) %>% 
  head(10)

df_plot %>% 
  filter(horizon == 2) %>% 
  select(mnemonic, unconditional, horizon) %>% 
  pivot_wider(names_from = "horizon", values_from = "unconditional") %>% 
  arrange(`2`) %>% 
  head(10)
```

```{r}
df_plot %>% 
  mutate(diff_fore = conditional_hard - unconditional) %>% 
  group_by(category, horizon) %>% 
  summarise(mean_diff = mean(diff_fore)) %>% 
  pivot_wider(values_from = mean_diff, names_from = horizon)
```
#### Results

> Figure XX illustrates the results for point forecasts. Turning to the peformance of the unconditional forecasts first, we see that for all horizons and the majority of variables, the factor model produces forecasts that perform similar to the autoregressive benchmark with relative RMSFE between 0.9 and 1.1. Noteworthy outliers in this context are the two price series - XX and YY- at $h=0$ where the performanc is considerably worse than the benchmark. Conversely, for a few series there are noticeable gains over the simple univariate AR(2) model and these persist or even increase for larger horizons. For example, for $h=0$ the relative RMSFE for the construction price index and gross value added in the industrial sector is around 0.6. Smaller but still noteworthy improvements can also be found for two series from the expenditure side of the national accounts: equipment investment as well as the deflator for residential investment. For larger $h$, these relative gains even increase: the relative RMSFE for the construction subcomponent of the PPI reaches 0.35; for gross value added in the industrial sector is 0.4. In addition to these variables, at $h=2$ there also noteworthy improvements for series that are likely highly correlated to gross industrial value added like production in the industrial and construction sector as well as industrial turnover. Note, however, that these improvements are relative to the benchmark, the performance of which considerably worsens as the forecast horizon increases. Indeed, for all series discussed above, the (absolute) RMSFE increases in $h$. Focusing on how the forecast performance increases when we condition on professional forecasters' views on GDP growth and CPI inflation, we see that there are only large differences in the relative RMSFE between the unconditional and conditional forecasts for $h=0$. Perhaps not surprisingly gains to conditioning  arise for those series for which the models already produce decent forecasts (relative to autoregressive benchmarks) and are mainly concentrated to series from the national accounts group and - to some extent - indicators capturing real activity (but not part of the national accounts). For the labor market indicators and CPI and PPI inflation, we find much smaller gains from conditioning.  

#### Notes for figure

> The figure shows the root mean squared forecast errors (RMSFE) r corresponding to unconditional forecasts (x-axis) and forecasts conditional on professional forecasters' view on GDP growth and CPI inflation (y-axis) for different time series. For each series, the RMSFE is relative to an autoregressive benchmark. For details, see the main text.Entries above (below) the 45-degree line indicate that conditional forecasts perform worse (better) than the unconditional ones.  

### Density forecast accuracy

```{r}
# df_plot
str_title <- "rel. CRPS"
df_eval %>%
  filter(horizon >= 0) %>%
  pivot_longer(c(rel_rmsfe, rel_crps, rel_logs), names_to = "score", values_to = "value") %>%
  filter(value != Inf, score == "rel_crps") %>%
  pivot_wider(names_from = type, values_from = value) -> df_plot 

# data-dependent lims!
ylims <- c(min(min(df_plot$unconditional), min(df_plot$conditional_hard)),
           max(max(df_plot$unconditional), max(df_plot$conditional_hard)))
xlims <- ylims

# manually set lims, excluding "outliers"
# ylims <- c(0.8, 1.2)
# xlims <- ylims

ggplot(df_plot, aes(x = unconditional, 
                    y = conditional_hard, 
                    color = category)
       )+
  geom_point(size = 2)+
  geom_abline()+
  geom_vline(xintercept = 0,  size = 0.2)+
  geom_hline(yintercept = 0, size = 0.2)+
  facet_wrap(~horizon_as_factor(horizon), nrow = 1)+
  xlim(xlims)+
  ylim(ylims)+
  scale_color_jco()+
  #theme_minimal()+
  theme(legend.position="bottom", 
        legend.text = element_text(size = 8),
        legend.title = element_blank())+
  labs(title = "", x = "unconditional forecasts", y = "conditional forecasts")
```

#### Export plot to pdf

```{r}
ggsave(filename = "../figures/fig_eval_crps.pdf", 
       width = 10, height = 6, units = "in")
```

#### Analyis

```{r}
df_plot %>% 
  filter(horizon == 0) %>% 
  select(mnemonic, unconditional, horizon) %>% 
  pivot_wider(names_from = "horizon", values_from = "unconditional") %>% 
  arrange(`0`) %>% 
  head(10)

df_plot %>% 
  filter(horizon == 1) %>% 
  select(mnemonic, unconditional, horizon) %>% 
  pivot_wider(names_from = "horizon", values_from = "unconditional") %>% 
  arrange(`1`) %>% 
  head(10)

df_plot %>% 
  filter(horizon == 2) %>% 
  select(mnemonic, unconditional, horizon) %>% 
  pivot_wider(names_from = "horizon", values_from = "unconditional") %>% 
  arrange(`2`) %>% 
  head(10)
```


#### Results

> Regarding density forecast accuracy, the overall results are very similar (Figure XX). The gains in unconditional forecast accuracy are less pronounced at the extremes for $h=2$ but nevertheless sizeable. In terms of the gains from conditioning the results are virtually identical to those obtained for point forecasts. 


#### Notes for figure

> The figure shows the continous rank probability score (CRPS) corresponding to unconditional forecasts (x-axis) and forecasts conditional on professional forecasters' view on GDP growth and CPI inflation (y-axis) for different time series. For each series, the CRPS is relative to an autoregressive benchmark. For details, see the main text. Entries above (below) the 45-degree line indicate that conditional forecasts perform worse (better) than the unconditional ones.  

### National accounts series


```{r}
# df_plot
df_eval %>%
  filter(horizon >= 0, 
         category %in% c("national accounts (expenditure)", "national accounts (production)")) %>%
  pivot_longer(c(rel_rmsfe, rel_crps, rel_logs), names_to = "score", values_to = "value") %>%
  filter(value != Inf, score == "rel_rmsfe") %>%
  pivot_wider(names_from = type, values_from = value) -> df_plot 

# data-dependent lims!
ylims <- c(min(min(df_plot$unconditional), min(df_plot$conditional_hard)),
           max(max(df_plot$unconditional), max(df_plot$conditional_hard)))
xlims <- ylims

df_plot[df_plot$category == "national accounts (expenditure)", "category"] <- "expenditure"
df_plot[df_plot$category == "national accounts (production)", "category"] <- "production"

ind_prices <- substr(df_plot$mnemonic, 1, 1) == "p"

df_plot$type <- "volume"
df_plot$type[ind_prices] <- "prices"

ggplot(df_plot, aes(x = unconditional, 
                    y = conditional_hard, 
                    color = category,
                    shape = type)
       )+
  geom_point(size = 2)+
  geom_abline()+
  #geom_vline(xintercept = 1.0,  size = 0.2)+
  #geom_hline(yintercept = 1.0, size = 0.2)+
  facet_wrap(~horizon_as_factor(horizon), nrow = 1)+
  xlim(xlims)+
  ylim(ylims)+
  scale_color_jco()+
  #theme_minimal()+
  theme(legend.position="bottom", 
        legend.text = element_text(size = 8),
        legend.title = element_blank())+
  labs(title = "", x = "unconditional forecasts", y = "conditional forecasts")
```
#### Export plot to pdf

```{r}
ggsave(filename = "../figures/fig_eval_natacc.pdf", 
       width = 10, height = 6, units = "in")
```

#### Analyis

```{r}
df_plot %>% mutate(diff_fore = conditional_hard-unconditional) %>% group_by(type, horizon) %>% summarise(mean(diff_fore))
```

#### Results

> The results for the national accounts merit special attention as they are arguably the series which analysts are most important to analysts. In addition, since our dataset includes both chain indices that capture price-adjusted or volume changes in consumption, investment etc. as well as the corresponding deflators, it is of interest to see if the benefits from conditioning differ across the two types of variables. On average, there are indeed sizeable differences in the reduction of the RMSFE of the conditional forecasts for quantity and price variables. For $h=0$, the RMSFE drops by roughly 0.1 for the former; for the latter the gains from conditioning are much less pronounced (-0.02). As the forecast horizons increases, the reduction in the RMSFE for the volume series drops considerably but still amounts to close to four percent on average. Conditional forecasts of the deflators for $h=\{1, 2\}$ are about as accurate (relative to unconditional forecasts) on average for higher forecasts. Figure XX plots the results for all national accounts series. 

#### Notes for figure

> The figure shows the root mean squared forecast error (RMSFE) corresponding to unconditional forecasts (x-axis) and forecasts conditional on professional forecasters' view on GDP growth and CPI inflation (y-axis) for the national accounts series. For each series, the RMSFE is relative to an autoregressive benchmark. For details, see the main text. Entries above (below) the 45-degree line indicate that conditional forecasts perform worse (better) than the unconditional ones.  


## Robustness

### Different subsamples

```{r}
str_model <- "Nr2_Nj1_Np2_Ns0"
subsample <- c(as.Date("2011-01-01"), max(df_eval_models$quarter))

df_eval <- calc_relative_forecast_accuracy(df_eval_models, 
                                           df_eval_benchmark, 
                                           str_model, 
                                           subsample)

```

```{r}
# df_plot
str_title <- "post-crisis sample: RMSFE and CRPS"
df_eval %>%
  filter(horizon >= 0) %>%
  pivot_longer(c(rel_rmsfe, rel_crps, rel_logs), names_to = "score", values_to = "value") %>%
  filter(value != Inf, score %in% c("rel_rmsfe", "rel_crps")) %>%
  pivot_wider(names_from = type, values_from = value) -> df_plot 

# data-dependent lims!
ylims <- c(min(min(df_plot$unconditional), min(df_plot$conditional_hard)),
           max(max(df_plot$unconditional), max(df_plot$conditional_hard)))
xlims <- ylims

# manually set lims, excluding "outliers"
# ylims <- c(0.8, 1.2)
# xlims <- ylims

ggplot(df_plot, aes(x = unconditional, 
                    y = conditional_hard, 
                    color = category,
                    shape = score)
       )+
  geom_point(size = 2, alpha = 0.7)+
  geom_abline()+
  geom_vline(xintercept = 0,  size = 0.2)+
  geom_hline(yintercept = 0, size = 0.2)+
  facet_wrap(~horizon_as_factor(horizon), nrow = 1)+
  xlim(xlims)+
  ylim(ylims)+
  scale_color_jco()+
  #theme_minimal()+
  theme(legend.position="bottom", 
        legend.text = element_text(size = 8),
        legend.title = element_blank())+
  labs(title = str_title, x = "unconditional forecasts", y = "conditional forecasts")
```

### Different models

Compare the models' performance when $R\neq 2$
```{r}
subsample <- c(min(df_eval_models$quarter), max(df_eval_models$quarter))
df_compare_R <- data.frame()

for (nr in c(2, 1, 5, 8))
{
  str_model <- paste0("Nr", nr, "_Nj1_Np2_Ns0")

  calc_relative_forecast_accuracy(df_eval_models, 
                               df_eval_benchmark, 
                               str_model, 
                               subsample) %>% 
  mutate(Nr = paste0(nr)) %>% 
  rbind(df_compare_R) -> df_compare_R
}
```
```{r}
df_compare_R %>% 
  filter(horizon >= 0) %>%
  pivot_longer(c(rel_rmsfe, rel_crps, rel_logs), names_to = "score", values_to = "value") %>%
  filter(value != Inf, score %in% c("rel_rmsfe")) %>%
  select(-category, -score) %>%
  pivot_wider(id_cols = c(mnemonic, type, horizon), names_from = "Nr", values_from = "value")%>%
  rename(base_model = `2`) %>%
  pivot_longer(cols = c(`1`, `5`, `8`), 
               names_to = "Nr", 
               values_to = "other_models") -> df_plot

# data-dependent lims!
ylims <- c(min(min(df_plot$base_model), min(df_plot$other_models)),
           max(max(df_plot$base_model), max(df_plot$other_models)))
xlims <- ylims


ggplot(df_plot, aes(x = base_model, y = other_models, color = Nr, shape = type))+
  geom_point(size = 2, alpha = 0.7)+
  geom_abline()+
  geom_vline(xintercept = 0,  size = 0.2)+
  geom_hline(yintercept = 0, size = 0.2)+
  facet_wrap(~horizon_as_factor(horizon), nrow = 1)+
  scale_x_continuous(limits = xlims, 
                     breaks = seq(0.5, 1.5, by = 0.5))+
  scale_y_continuous(limits = ylims, 
                     breaks = seq(0.5, 1.5, by = 0.5))+
  scale_shape_discrete(name = element_blank())+
  scale_color_jco(name = "R")+
  theme(legend.position="bottom", 
        legend.text = element_text(size = 8))+
  guides(shape = guide_legend(order = 2),col = guide_legend(order = 1))+
  labs(title = "", x = "rel. RMSFE, R=2", 
       y = paste0("rel. RMSFE, R≠2"))
```

#### Analysis

For which series do the forecasts improve at $h=0$? 

```{r}
df_plot %>% 
  filter(Nr == 8) %>% 
  mutate(diff_models = base_model - other_models) %>% 
  arrange(desc(diff_models)) %>% 
  select(mnemonic, type, diff_models, horizon) %>% 
  head(20)
```


#### Results

> Besides the evaluation period, the results presented above may also be sensitive to the model specification. In particular, they were obtained under a given number of factors, $R=2$. However, differently specified models yield similar forecast performance. Figure XX shows the relative RMSFE of the model with $R=2$ on which the results above are based as well as both smaller ($R=1$) and larger ($R = {5,8}$) models. There is little indication that model specification systematically alters the forecast performance. Conditional forecasts at $h=0$ for the change in the producer index improve substantially when $R=8$ with the relative RMSFE decreasing by 0.25. To a lesser extent this also holds for the unconditional forecasts with the relative RMSFE dropping by around 0.1. However, this is the exception and for other horizons and variables we find much smaller differences in either way. Consequently, the points corresponding to the relative RMSFE for $R={1, 5, 8}$ in Figure XX all hug the 45-degree line, which indicates identical forecast performance, quite closely. 

#### Notes for figure

> Note: The figure shows the root mean squared forecast error (RMSFE) relative to the autoregressive benchmarks for different number of factors. On the x-axis are the values of the RMSFE for the model with 2 factors on which the main results are based. On the y-axis are the RMSFE for alternative model specifications, differentiated by colors. Entries above (below) the 45-degree line indicate that the model with 2 factors performs better (worse) than the alternative models.  

#### Export plot to pdf

```{r}
ggsave(filename = "../figures/fig_eval_robustness_Nr.pdf", 
       width = 10, height = 6, units = "in", device=cairo_pdf)
```



