---
title: "Simulation results"
output:
  html_document:
    df_print: paged
---

# Overview

This notebook loads the output of the runtime simulations and transform and analyzes the output. Also produces a plot and calculates results that are used in the paper. 

## Clear workspace, load packages

```{r}
rm(list = ls())
setwd("C:/Users/Philipp/Documents/GitHub/condfcast-precsampler/sim")
dir_fig <- "C:/Users/Philipp/Documents/GitHub/condfcast-precsampler/figures/"
library(ggplot2)
library(dplyr)
library(tidyr)
library(scales)
library(ggsci)
```


## Set-up

```{r}
#dir_in <- "./../../../Dissertation/condfcast-precsampler/sim/out/"
dir_in <- "./out local/"
file_prefix <- "runtime"
Ng <- 10
Nm <- 1000
models <- c(rep("ssm", 3), rep("var", 3))
Nhs <- c(5, 20, 50)
Nconds <- c(10, 50, 75)

dims = c("Nn_20_Ns_2",
         "Nn_100_Ns_2",
         "Nn_100_Ns_10",
         "Nn_3_Np_4",
         "Nn_20_Np_4",
         "Nn_100_Np_4")

ftypes = c("uncond", "cond_hard", "cond_soft")

samplers <- c("CK", "DK", "HS")
```

Names for factor levels

```{r}
names_samplers <- c("Carter and Kohn (1994)", "Durbin and Koopman (2002)", "precision-sampler")
names_ftypes <- c("unconditional forecasts", "hard conditioning", "soft conditioning")
names_dims = c("small factor model",
               "large N factor model",
               "large factor model",
               "small-sized VAR",
               "medium-sized VAR",
               "large-sized VAR")
```


## Load simulation output

```{r}
dat_raw <- data.frame() # initialize empty data frame

for (g in seq(1, Ng)){
  for (s in samplers){
    for (f in ftypes){
      for (h in Nhs){
        for (ncond in Nconds){
          for (d in seq(1, length(dims))){
            filename <- paste(file_prefix, s, f, 
                              models[d], dims[d], 
                              "Nh", h, "Ncond", ncond, "g", g, sep = "_")
            tmp <- read.csv(paste0(dir_in, filename, ".csv"), header = FALSE)
            dat_raw <- rbind(dat_raw, data.frame(value = tmp[, 1], sim = g,
                                         sampler = s, ftype = f, mod = dims[d],
                                         Nh = paste(h), Ncond = paste(ncond/100)
                                         )
                         )
          }
        }
      }
    }
  }
}
```

## Transform output

```{r}
# Calculate relative run time (precision-sampler = 1)
dat_raw %>% 
pivot_wider(names_from = sampler, 
            values_from = value) %>%
  mutate(CK = CK / HS,
         DK = DK / HS,
         HS = HS / HS) %>%
  pivot_longer(cols = -c(sim, ftype, mod, Nh, Ncond), names_to = "sampler", values_to = "value_rel") -> dat

dat <- merge(dat, dat_raw, by = intersect(names(dat), names(dat_raw)))

# convert strings to factors----
dat$mod <- factor(dat$mod, levels = dims, labels = names_dims)
dat$ftype <- factor(dat$ftype, levels = ftypes, labels = names_ftypes)
dat$sampler <- factor(dat$sampler, levels = samplers, labels = names_samplers)
dat$Nh <- factor(dat$Nh, levels = Nhs, labels = paste0("H=", Nhs))
dat$Ncond <- factor(dat$Ncond, levels = paste0(Nconds/100), labels = paste0("N_c=", Nconds/100))

```

## Analyze output

Calculate mean and standard deviation of runtime to see if there are any outliers and if the runtime is always above $0.1s$ as suggested by the [Matlab documentation]{https://de.mathworks.com/help/matlab/matlab_prog/measure-performance-of-your-program.html}

```{r}
dat %>% 
  group_by(mod, sampler, ftype, Ncond, Nh) %>%
  summarise(mean_runtime = mean(value),
            sd_runtime = sd(value)) %>%
  ungroup() -> dat_tmp
```

```{r}
dat_tmp %>% ggplot(aes(x = mod, y = mean_runtime, color = sampler, shape = ftype))+
    geom_point(position = position_dodge(width = 1.0, 
                                           preserve = c("total")
                                           ))+
  facet_grid(Ncond ~ Nh)+
  theme(text = element_text(size=7),
        axis.text.x = element_text(angle=90, hjust = 1),
        legend.position="right")
```


```{r}
dat_tmp %>% ggplot(aes(x = mod, y = sd_runtime, color = sampler, shape = ftype))+
    geom_point(position = position_dodge(width = 1.0, 
                                           preserve = c("total")
                                           ))+
  facet_grid(Ncond ~ Nh)+
  theme(text = element_text(size=7),
        axis.text.x = element_text(angle=90, hjust = 1),
        legend.position="right")
```

Is there a relationship between average runtime and standard deviation?

```{r}
dat_tmp %>% ggplot(aes(x = mean_runtime, y = sd_runtime, color = interaction(sampler, mod), shape = ftype))+
    geom_point()+
  facet_grid(Ncond ~ Nh)+
  theme(text = element_text(size=7),
        axis.text.x = element_text(angle=90, hjust = 1),
        legend.position="right")
```


```{r}
filter(dat, value < 0.1)
```


```{r, fig.width = 10, fig.height=8}
dat %>% group_by(mod, sampler, ftype, Ncond, Nh) %>%
  mutate(upper = max(value),
         lower = min(value)) %>%
  ungroup() %>%
  ggplot(aes(x = mod, y = value, color = sampler, linetype = ftype))+
  geom_linerange(aes(ymin = lower, ymax = upper), 
                 position = position_dodge(width = 0.9, 
                                           preserve = c("total")
                                           )
                 )+
  facet_grid(Ncond~Nh)+
  scale_y_continuous(trans='log10')+
  theme(text = element_text(size=9),
        axis.text.x = element_text(angle=90, hjust = 1),
        legend.position="top",
        axis.title.x = element_blank(), axis.title.y = element_blank(),
        panel.grid.minor.y = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.major.x = element_blank(),)
```

```{r, fig.width = 10, fig.height=8}
dat %>% filter(ftype == "hard conditioning") %>%
  ggplot(aes(x = mod, y = value, color = sampler))+
  geom_boxplot()+
  facet_grid(Ncond~Nh)+
  scale_y_continuous(trans='log10')+
  theme(text = element_text(size=9),
        axis.text.x = element_text(angle=90, hjust = 1),
        legend.position="top",
        axis.title.x = element_blank(), axis.title.y = element_blank(),
        panel.grid.minor.y = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.major.x = element_blank(),)
```



```{r, fig.width = 10, fig.height=8}
dat %>% filter(ftype == "soft conditioning") %>%
  ggplot(aes(x = mod, y = value, color = sampler))+
  geom_boxplot()+
  facet_grid(Ncond~Nh)+
  scale_y_continuous(trans='log10')+
  theme(text = element_text(size=9),
        axis.text.x = element_text(angle=90, hjust = 1),
        legend.position="top",
        axis.title.x = element_blank(), axis.title.y = element_blank(),
        panel.grid.minor.y = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.major.x = element_blank(),)
```



## $g=1$

```{r, fig.width = 10, fig.height=10}
dat %>% 
  filter(sim == 1) %>%
  ggplot(aes(x = mod, y = value, color = sampler, shape = ftype, group = sampler))+
  geom_point(position = position_dodge(width = 0.9, 
                                       preserve = c("total")
                                       ))+
  facet_grid(Nh~Ncond)+
  scale_y_continuous(trans='log10')+
  scale_shape(solid = FALSE)+
  theme(text = element_text(size=9),
        axis.text.x = element_text(angle=90, hjust = 1),
        legend.position="top",
        axis.title.x = element_blank(), axis.title.y = element_blank(),
        panel.grid.minor.y = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.major.x = element_blank(),)
```


# Tidy up plot for paper

```{r,fig.width = 10, fig.height=10}
dat %>% 
  filter(sim == 1,
         !(sampler == "precision-sampler"),
         !(ftype == "unconditional forecasts"),
         !(mod %in% c("large N factor model", "small-sized VAR")),
         Ncond == "N_c=0.1") %>%
  ggplot(aes(x = mod, y = value_rel, 
             color = sampler, shape = ftype, group = sampler))+
  geom_point(size = 4, stroke = 2, position = position_dodge(width = 0.3, 
                                              preserve = c("total")
                                              ))+
  geom_hline(yintercept = 1)+
  guides(color = guide_legend(order = 1), 
         shape = guide_legend(order = 2))+
  scale_shape(solid = FALSE)+
  scale_color_jco()+
  scale_y_continuous(trans='log10', breaks = c(1, 2, 10, 50))+
  facet_wrap(~Nh, ncol = 1)+
  theme_bw()+
  theme(text = element_text(size=15),
        legend.position="top", legend.title = element_blank(),
        panel.grid.minor.y = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.major.x = element_blank(),
        axis.text.x = element_text(angle=30, hjust = 1),
        axis.title.x = element_blank(),
        axis.title.y = element_blank()) 
```

```{r,fig.width = 10, fig.height=10}
dat %>% 
  filter(!(sampler == "precision-sampler"),
         !(ftype %in% c("unconditional forecasts", "soft conditioning")),
         !(mod %in% c("large N factor model", "small-sized VAR")),
         Ncond == "N_c=0.1") %>%
  ggplot(aes(x = mod, y = value_rel, 
             color = sampler))+
  geom_boxplot()+
  guides(color = guide_legend(order = 1), 
         shape = guide_legend(order = 2))+
  scale_shape(solid = FALSE)+
  scale_color_jco()+
  scale_y_continuous(trans='log10', breaks = c(1, 2, 10, 50), limits = c(0.7, 20))+
  facet_wrap(~Nh, ncol = 1)+
  theme_bw()+
  theme(text = element_text(size=15),
        legend.position="top", legend.title = element_blank(),
        panel.grid.minor.y = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.major.x = element_blank(),
        axis.text.x = element_text(angle=30, hjust = 1),
        axis.title.x = element_blank(),
        axis.title.y = element_blank()) 
```

```{r,fig.width = 10, fig.height=10}
dat %>% 
  filter(!(sampler == "precision-sampler"),
         !(ftype %in% c("unconditional forecasts", "hard conditioning")),
         !(mod %in% c("large N factor model", "small-sized VAR")),
         Ncond == "N_c=0.1") %>%
  ggplot(aes(x = mod, y = value_rel, 
             color = sampler))+
  geom_boxplot()+
  guides(color = guide_legend(order = 1), 
         shape = guide_legend(order = 2))+
  scale_shape(solid = FALSE)+
  scale_color_jco()+
  scale_y_continuous(trans='log10', breaks = c(1, 2, 10, 50), limits = c(1, 90))+
  facet_wrap(~Nh, ncol = 1)+
  theme_bw()+
  theme(text = element_text(size=15),
        legend.position="top", legend.title = element_blank(),
        panel.grid.minor.y = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.major.x = element_blank(),
        axis.text.x = element_text(angle=30, hjust = 1),
        axis.title.x = element_blank(),
        axis.title.y = element_blank()) 
```


```{r}
#ggsave(paste0(dir_fig, "fig_simulations_Ncond_10.pdf"), width = 10, height = 10, units = "in")
```


