---
title: "Reuters Poll forecasts for GDP and CPI growth"
output: html_notebook
---

## Set-up
```{r}
# clear workspace
rm(list = ls())

# working directory
setwd("C:/Users/Philipp/Documents/GitHub/condfcast-precsampler/data")

# packages
library(dplyr)
library(lubridate)
library(tidyr)
library(ggplot2)
library(bundesbank)
library(scales)
library(ggsci)

# directories
dir_forecasts <- "/raw/Reuters Poll/"

# evaluation period
yys <- 2006:2018 # years
qqs <- 1:4 # quarters

# max and min horizons
max_h <- 3 # 3Q-ahead forecast (relative to the quarter in which the forecast is made)
min_h <- -1 # backcast
```

## Functions

```{r}
# define helper function
dropNA_first <- function(df)
# function that removes NA observations and drops first element
# output is a vector
{
  tmp <- df[!is.na(df)] # rem NA
  out <- tmp[-1] # drop first element
}
```

Source some other functions

```{r}
source("../functions/functions.R")
```


## Extract Reuters Poll data from raw csv-files

### GDP

Function to read in Reuters Poll forecasts
```{r}
get_gdp_forecasts <- function(yy, qq, filename, max_h, min_h)
{
  # read in csv
  
  if (file.exists(filename))
  {
    dat <- read.csv(filename)
  } else
  {
    break
  }

  
  # get row indices 
  ind_median <- grep("median", dat[, 1], ignore.case = TRUE)
  ind_min <- grep("min", dat[, 1], ignore.case = TRUE)
  ind_max <- grep("max", dat[, 1], ignore.case = TRUE)
  ind_date <- grep("consensus", dat[, 1], ignore.case = TRUE)

  # extract values
  dates_fore_tmp <- dmy(dropNA_first(dat[ind_date, ]))
  medf_tmp <- as.numeric(dropNA_first(dat[ind_median, ]))
  minf_tmp <- as.numeric(dropNA_first(dat[ind_min, ]))
  maxf_tmp <- as.numeric(dropNA_first(dat[ind_max, ]))
  quarter_date <- make_date(year = yy, month = 3 * qq - 2, day = 1) 
  
  # calculate forecast horizon (in quarters)
  horizon_tmp <- round(as.numeric((quarter_date - dates_fore_tmp) / 90), digits = 0)
  
  # insert NA where no values are present between max_h and min_h
  seq_h <- seq(max(c(max(horizon_tmp), max_h)), min(c(min(horizon_tmp), min_h)))
  ind_h <- seq_h %in% horizon_tmp
  tmp_NA <- rep(NA, length(seq_h))
  medf <- tmp_NA; medf[which(ind_h)] <- medf_tmp
  minf <- tmp_NA; minf[which(ind_h)] <- minf_tmp
  maxf <- tmp_NA; maxf[which(ind_h)] <- maxf_tmp
  dates_fore <- tmp_NA; dates_fore[which(ind_h)] <- dates_fore_tmp
  dates_fore <- as_date(dates_fore)
  horizon <- seq_h

  # store in df
  df <- data.frame(dates_fore = as_date(dates_fore), 
                   med = medf,
                   min = minf,
                   max = maxf,
                   quarter = quarter_date,
                   horizon = horizon
                  )
  
 return(df)
}
```


```{r}
dir_in <- paste0(dir_forecasts, "GDP/")
df_gdp <- data.frame() 
for (yy in yys)
{
  for (qq in qqs)
  {
    filename <- paste0(getwd(), dir_in, yy, "Q", qq, ".csv")
    if (file.exists(filename))
      df_gdp <- rbind(df_gdp, get_gdp_forecasts(yy, qq, filename, max_h, min_h))
    else
      next
  }
}
```


### CPI

The original Reuters Poll reports professional forecasters' CPI forecasts in terms of y/y changes and need to be converted to q/q growth rates given the available data at the time

To this end

- download the corresponding vintage of data available at the time the forecasts were made using the package `bundesbank`
- given the CPI data, reconstruct the (log) level of the CPI $log(cpi)_{t+h}$ given the forecasts $\Delta_4 cpi^f_{t+h|t}$, i.e. $\Delta_4 cpi_{t} = log(cpi)_{t} - log(cpi)_{t-4}$ 
$$
log(cpi)_{t+h} = log(cpi)_{t+h-4} + \Delta_4 cpi^f_{t+h|t}
$$
- calculate the corresponding q/q growth rates and take the min, max and median for each quarter

#### Functions 

```{r}
get_individual_fore <- function(yy, qq, filename)
{
  
  dat <- read.csv(filename)
  
  ind_date <- grep("consensus", dat[, 1], ignore.case = TRUE)
  dates_fore_tmp <- dmy(dropNA_first(dat[ind_date, ]))
  
  tmp <- grep("Contributor", dat[, 1], ignore.case = TRUE)
  ind_contribs <- tmp + seq(1, nrow(dat)-tmp)
  
  df_fore_yoy <- data.frame()
  for (i in ind_contribs)
  {
    fore <- as.numeric(dropNA_first(dat[i, ]))
    name_contrib <- dat[i, 1]
    df_fore_yoy <- rbind(df_fore_yoy, 
                         data.frame(yoy = fore, 
                                    dates = dates_fore_tmp,
                                    name = name_contrib)
    )
    
  }
  
  df_fore_yoy$quarter_date <- make_date(year = yy, month = 3 * qq -2, day = 1) 
  
  df_fore_yoy$horizon <- round(as.numeric((df_fore_yoy$quarter_date - df_fore_yoy$dates) / 90), digits = 0)
  return(df_fore_yoy)
}
```

```{r}
transform_yoy_to_qoq <- function(cpi, df)
  {
    df %>%
      drop_na() %>%
      filter(dates == v) -> df
    
    unique_names <- unique(df$name)
    
    df_out <- data.frame()
    
    for (n in seq(1, length(unique_names)))
    {
    
      df %>%
        filter(name == unique_names[n]) %>%
        arrange(quarter_date) -> df_tmp
      
      cpi_tmp <- cpi
      cpi_tmp$log <- log(cpi_tmp$raw)
      
      for (i in seq(1, nrow(df_tmp)))
      {
        ind_row <- which(cpi_tmp$date == df_tmp$quarter_date[i])
        cpi_tmp$log[ind_row]<- df_tmp$yoy[i]/100 + cpi_tmp$log[ind_row - 4]
      }
      
      cpi_tmp$qoq <- c(NA, 100 * diff(cpi_tmp$log))
  
      df_out <- rbind(df_out, 
                      merge(df_tmp, 
                            select(cpi_tmp, date, qoq), 
                            by.x = "quarter_date", by.y = "date")
                      )
      df_out$qoq[is.nan(df_out$qoq)] <- NA
    }
    
    return(df_out)
}
```


#### Read in individual forecasts from csv's

```{r, message = F, warning = F}
dir_in <- "/raw/Reuters Poll/CPI/"
df_cpi_yoy <- data.frame()
for (yy in yys)
{
  for (qq in qqs)
  {
    filename <- paste0(getwd(), dir_in, yy, "Q", qq, ".csv")
    if (file.exists(filename))
      df_cpi_yoy <- rbind(df_cpi_yoy, get_individual_fore(yy, qq, filename))
    else
      next
  }
}

df_cpi_yoy$quarter_date <- make_date(
                            year = year(df_cpi_yoy$quarter_date),
                            month = month(df_cpi_yoy$quarter_date) + 2,
                            day = 1L
                                    )
```


#### Transform to q/q growth rates

```{r, message = F, warning = F}
vintages <- unique(df_cpi_yoy$dates)
df_cpi_qoq <- data.frame()
for (v in vintages)
{
  download_realtime_data("BBKRT.M.DE.Y.P.PC1.PC100.R.I", v) %>% 
    aggregate_to_Q("M") -> cpi
  
  df_cpi_qoq <- rbind(df_cpi_qoq, transform_yoy_to_qoq(cpi, df_cpi_yoy))
}


df_cpi_qoq %>% 
 drop_na %>% 
 group_by(quarter_date, horizon, dates) %>% 
 summarise(min = min(qoq, na.rm = T), 
           max = max(qoq, na.rm = T), 
           med = median(qoq, na.rm = T)) %>%
 select(dates_fore = dates, 
        med, min, max, 
        quarter = quarter_date, horizon) -> df_cpi
```


### Plots

Combine datasets into `df_plot`
```{r}
df_cpi$var <- "cpi"

df_gdp$var <- "gdp"

df_plot <- rbind(df_cpi, df_gdp)
```

```{r}
# change labels of variables
df_plot$var[df_plot$var == "cpi"] <- "CPI inflation"
df_plot$var[df_plot$var == "gdp"] <- "GDP growth"

# filter horizons, then convert to factor
df_plot %>%
  filter(horizon <=2) %>%
  filter(!(var == "cpi" & horizon == -1)) -> df_plot # also delete h=-1 for cpi!

df_plot$horizon <- factor(paste0("h=", df_plot$horizon))
```

```{r}
df_plot %>% 
  filter(horizon != "h=-1") %>% 
  ggplot(aes(x = quarter, y = med, group = horizon, color = horizon))+
  geom_line()+
  geom_point()+
  facet_wrap(~var, nrow = 2, scales = "free_y")+
  labs(x = "", y = "percent", caption = "")+
  scale_color_jco()+
  scale_y_continuous(labels = scales::number_format(accuracy = 0.1))+
  scale_x_date(date_breaks = "2 year", date_labels = "%Y")+
  theme_minimal()+
  theme(legend.title = element_blank())
```

```{r}
ggsave(paste0("../figures/fig_ReutersPoll.pdf"), width = 10, height = 5, units = "in")
```


## Export data

Need to export 

- the actual forecasts - median, min and max - as well as
- the list of vintages for which all forecasts are available 


### List of vintages

Determine the vintages that will be used in the forecast evaluation. For these we require that all nowcasts and forecasts (and backcasts in the case of GDP) are available. 

First, remove the backcasts for CPI. These are not necessary as we usually have a flash estimate of previous quarter CPI inflation and do not rely on it. Moreover, it appears to be missing during the start of the sample, so including it would mean that we discard a large part of the potential evaluation sample. 

```{r}
df_cpi <- filter(df_cpi, horizon != -1) 
```

Find the intersection of vintages for GDP and CPI

```{r}
vint_cpi <- as.character(unique(df_cpi$dates_fore))
vint_gdp <- as.character(unique(df_gdp$dates_fore))

vint_cpi <- vint_cpi[!is.na(vint_cpi)]
vint_gdp <- vint_gdp[!is.na(vint_gdp)]
vintages <- intersect(vint_cpi, vint_gdp)
```

Check that for these vintages we have all available forecasts, i.e. no missing ones in between!

```{r}
h_start_gdp <- -1
h_start_cpi <- 0
h_max <- 2
counter <- 1
complete_vintage <- matrix(FALSE, nrow = length(vintages), ncol = 2) # col1 = cpi, col2 = gdp
for (v in vintages)
{
  df_cpi %>% filter(dates_fore == v) -> df_tmp
  
  if (nrow(df_tmp) != 0)
  {
    #diff_cpi <- setdiff(seq(h_start_cpi, max(h_max, max(df_tmp$horizon))), df_tmp$horizon)
    df_tmp <- filter(df_tmp, horizon >=h_start_cpi, horizon <= h_max)
    #if (length(diff_cpi) == 0)
    if (!any(is.na(df_tmp$med)) & nrow(df_tmp) == h_max - h_start_cpi + 1)
      complete_vintage[counter, 1] = TRUE
  }
  
  df_gdp %>% filter(dates_fore == v) -> df_tmp
  if (nrow(df_tmp) != 0)
  {
    #diff_gdp <- setdiff(seq(h_start_gdp, max(h_max, max(df_tmp$horizon))), df_tmp$horizon)
    df_tmp <- filter(df_tmp, horizon >=h_start_gdp, horizon <= h_max)
    #if (length(diff_gdp) == 0)
    if (!any(is.na(df_tmp$med)) & nrow(df_tmp) == h_max - h_start_gdp + 1)
      complete_vintage[counter, 2] = TRUE
  }
  
  counter <- counter + 1
}

df_vintages <- data.frame(v = vintages, 
                          complete_cpi = complete_vintage[, 1],
                          complete_gdp = complete_vintage[, 2]
                          )

df_vintages <- filter(df_vintages, complete_cpi == T, complete_gdp == T)
```

Export this list as `list_vintages.csv`

```{r}
list_vintages <- df_vintages$v
write.table(file = "list_vintages.csv", list_vintages, row.names = F, col.names = F, quote = F)
```

### Forecasts 

```{r}
df_out <- rbind(df_gdp, df_cpi)

df_out <- filter(df_out, dates_fore %in% as_date(df_vintages$v))

save(file = "ReutersPoll_data.Rda", df_out)
```


